---
title: "List Column Switch"
output: 
  html_notebook: 
    toc: yes
    toc_float: true
---

## Summary

Convert a single column in a Synapse Table with comma-separated lists of values in `STRING` items into a `STRING_LIST` column (and save a copy of the original); remove the `STRING_LIST` column and revert to the original if you need to access the table with **synapser**.

## Setup

### Environment

```{r}
library(reticulate)
library(dccvalidator)
library(tidyverse)

# local conda env with Python 3.7 and synapseclient installed
use_condaenv("csbc-pson-dcc", required = TRUE)
```

### Synapse things

```{r}
synapseclient <- reticulate::import("synapseclient")
syntab <- reticulate::import("synapseclient.table")
syn <- synapseclient$Synapse()
syn$login()
```

## Functions

### Helpers

```{r, warning=FALSE}
#' Refresh the rows of a table in Synapse
#'
#' Note: this will overwrite all existing rows, not attempt to merge or
#' update only those rows with changes.
#'
#' @param table_id A string specifying the Synapse Table ID
#' @param update_df A dataframe or tbl representing the new version of the table
#' @param syn A reticulated synapseclient Python object
#' @param syntab A reticulated synapseclient.table module
#'
#' @return The reticulated synapseclient.table.CsvFileTable object representing
#' the updated table

#' @export
#'
#' @examples update_synapse_table("syn21980893", update_df, syn, syntab)
update_synapse_table <- function(table_id, update_df, syn, syntab) {
  current_rows <- syn$tableQuery(glue::glue("SELECT * FROM {table_id}"))
  syn$delete(current_rows)
  tmpfile <- fs::file_temp("rows.csv")
  write_csv(update_df, tmpfile)
  update_rows <- syntab$Table(table_id, tmpfile)
  syn$store(update_rows)
}

#' Convert comma-separated string values into a JSON array string
#'
#' @param str A string with values separated by commas
#'
#' @return The string containing a JSON array of the values
#' @export
#'
#' @examples
#' .csv_to_json("foo, bar, baz")
.delim_str_to_json <- function(str, delimiter = ",") {
  json_str <- str_split(str, delimiter)[[1]] %>% 
    str_trim() %>% 
    jsonlite::toJSON()
  if (json_str == "[null]") {
    "[]"
  } else {
    json_str
  }
}

#' Compute the maximum string length for a string list column with JSON arrays
#'
#' @param list_col A character vector with all items in the list column
#'
#' @return The (integer) length of the largest string
#' @export
#'
#' @examples 
#' .max_list_str_length(c('["123", "123456"]', '["1234", "12345"]' )
.max_list_str_length <- function(list_col) {
  map(list_col, function(l) {
  if (!is.na(l)) {
    jsonlite::fromJSON(l) %>% 
      map_int(str_length) %>% 
      max(na.rm = TRUE) %>%
      I
  } else {
    0L
  }
}) %>% 
    discard(is.infinite) %>%
    flatten_int() %>% 
    max()
}

.max_list_size <- function(list_col) {
  map(list_col, function(l) {
  if (!is.na(l)) {
    jsonlite::fromJSON(l) %>% 
      length()
  } else {
    0L
  }
}) %>% 
    discard(is.infinite) %>%
    flatten_int() %>% 
    max()
}
```

### The meat

#### `add_list_column`

```{r}
add_list_column <- function(table_id, column_name, delimiter = ",", 
                            syn, syntab) {
  df <- dccvalidator::get_synapse_table(table_id, syn)
  
  new_name <- str_c(column_name, "list", sep = "_")
  column_name <- rlang::ensym(column_name)
  new_name <- rlang::ensym(new_name)
  
  df <- df %>% 
    rowwise() %>% 
    mutate(!!new_name := .delim_str_to_json(!!column_name, delimiter)) %>% 
    ungroup()
  
  # df
  max_str_len <- .max_list_str_length(df[[rlang::as_string(new_name)]])*1.5
  max_str_len <- as.integer(round(max_str_len))
  print(max_str_len)
  max_size <- .max_list_size(df[[rlang::as_string(new_name)]]) + 1
  max_size <- as.integer(round(max_size))
  print(max_size)
  df
}
```

```{r}
add_list_column <- function(table_id, column_name, delimiter = ",", 
                            syn, syntab) {
  df <- dccvalidator::get_synapse_table(table_id, syn)
  
  old_name <- str_c(column_name, "asCsv", sep = "_")
  column_name <- rlang::ensym(column_name)
  old_name <- rlang::ensym(old_name)
  
  df <- df %>% 
    mutate(!!old_name := str_trim(!!column_name)) %>% 
    rowwise() %>% 
    mutate(!!column_name := .delim_str_to_json(!!column_name, delimiter)) %>% 
    ungroup()
  
  # df
  max_str_len <- .max_list_str_length(df[[rlang::as_string(column_name)]])*1.5
  max_str_len <- as.integer(round(max_str_len))
  print(max_str_len)
  df

  # col_id <-reticulate::iterate(syn$getTableColumns(table_id))  %>%
  #   map(~ jsonlite::fromJSON(.$json())) %>%
  #   keep(~ .$name == column_name) %>%
  #   map_chr("id")
  # old_column <- syn$getColumn(col_id)
  # 
  # 
  # new_column <- syn$store(
  #   syntab$Column(
  #     name = rlang::as_string(column_name),
  #     columnType = "STRING_LIST",
  #     maximumSize = max_str_len,
  #     facetType = "enumeration"
  #   )
  # )
  # 
  # schema <- syn$get(table_id)
  # schema$removeColumn(old_column)
  # schema$addColumn(new_column)
  # 
  # old_column_data <- jsonlite::fromJSON(old_column$json())
  # old_column <- syn$store(
  #   syntab$Column(
  #     name = rlang::as_string(old_name),
  #     columnType = old_column_data$columnType,
  #     maximumSize = old_column_data$maximumSize
  #   )
  # )
  # schema$addColumn(old_column)
  # 
  # schema <- syn$store(schema)
  # 
  # update_synapse_table(table_id, df, syn, syntab)
}
```

#### `remove_list_column`

```{r}
remove_list_column <- function(table_id, column_name, delimiter = ",", 
                               syn, syntab) {
  df <- dccvalidator::get_synapse_table(table_id, syn)
  
  new_name <- str_c(column_name, "asCsv", sep = "_")
  column_name <- rlang::ensym(column_name)
  new_name <- rlang::ensym(new_name)
  
  df <- df %>% 
    select(-!!column_name) %>% 
    rename(!!column_name := !!new_name)

  col_id <-reticulate::iterate(syn$getTableColumns(table_id))  %>% 
    map(~ jsonlite::fromJSON(.$json())) %>% 
    keep(~ .$name == column_name) %>% 
    map_chr("id")
  old_column <- syn$getColumn(col_id)
  
  new_col_id <-reticulate::iterate(syn$getTableColumns(table_id))  %>% 
    map(~ jsonlite::fromJSON(.$json())) %>% 
    keep(~ .$name == rlang::as_string(new_name)) %>% 
    map_chr("id")
  new_column <- syn$getColumn(new_col_id)
  
  schema <- syn$get(table_id)
  schema$removeColumn(old_column)
  schema$removeColumn(new_column)
  schema <- syn$store(schema)
  
  new_column_data <- jsonlite::fromJSON(new_column$json())
  new_column <- syn$store(
    syntab$Column(
      name = rlang::as_string(column_name),
      columnType = new_column_data$columnType,
      maximumSize = new_column_data$maximumSize,
      facetType = "enumeration"
    )
  )
  schema$addColumn(new_column)
  
  schema <- syn$store(schema)
  
  update_synapse_table(table_id, df, syn, syntab)
}
```


## Example

### Adding a list column

```{r}
table_id <- "syn21897968"
dccvalidator::get_synapse_table(table_id, syn)
```


```{r}
table_id %>% 
  add_list_column(column_name = "publication", ";", syn, syntab) %>% 
  select(publication)
```


```{r, warning=FALSE}
table <- add_list_column(
  "syn21980893", column_name = "list_col_1", ",", syn, syntab
)
```

```{r}
dccvalidator::get_synapse_table(table_id, syn)
```

### Removing a list column

```{r}
table <- remove_list_column(
  "syn21980893", column_name = "list_col_1", ",", syn, syntab
)
```


```{r}
dccvalidator::get_synapse_table(table_id, syn) 
```

## Fin.

```{r}
sessionInfo()
```

